# Data Carpentry - additional R exercises

## RNASeq data and annotation


2. Import the data from two tab-separated files into R. The
   [`kem_counts.tsv`](https://uclouvain-cbio.github.io/2019-05-20-UCLouvain/data/kem_counts.tsv)
   file contains RNA-Seq expression counts for 13 genes and 18 samples
   and
   [`kem_annot.tsv`](https://uclouvain-cbio.github.io/2019-05-20-UCLouvain/data/kem_annot.tsv)
   contains annotation about each sample. Read the data into two
   `tibbles` named `kem` and `annot` respectively and familiarise
   yourself with the content of the two new tables.
   
3. Convert the counts data into a long table format and annotation
   each sample using the experimental design.
   
4. Identity the three transcript identifiers that have the highest
   expression count over all samples.
   
5. Visualise the distribution of the expression for the three
   transcripts selected above in cell types A and B under both
   treatments.
   
6. For all genes, calculate the mean intensities in each experimental
   group (as defined by the `cell_type` and `treatment` variables).
   
7. Focusing only on the three most expressed transcripts and cell type
   A, calculate the fold-change induced by the treatment. The
   fold-change is the ratio between the average expressions in two
   conditions.
   
   
<details>
```{r}
library("tidyverse")

annot <- read_tsv("data/kem_annot.tsv")
kem <- read_tsv("data/kem_counts.tsv") %>%
    gather(key = sample_id, value = expression, -ref) %>%
    left_join(annot)
k <- kem %>%
    group_by(ref) %>%
    summarise(tot_exprs = sum(expression)) %>%
    arrange(desc(tot_exprs)) %>%
    select(ref) %>%
    head(3)
kem3 <- right_join(kem, k)
ggplot(kem3, aes(x = treatment, y = expression)) +
    geom_boxplot() +
    geom_jitter() +
    facet_grid(ref ~ cell_type)
kem %>%
    group_by(ref, cell_type, treatment) %>%
    summarise(mean_expression = mean(expression))
kem3 %>%
    filter(cell_type == "A") %>%
    group_by(ref, cell_type, treatment) %>%
    summarise(mean_expression = mean(expression)) %>%
    spread(key = treatment, value = mean_expression) %>%
    mutate(fold_change = stimulated/none)
```
</details>

## Quantitative proteomics statistical results

1. [Download](https://uclouvain-cbio.github.io/2019-05-20-UCLouvain/data/prot.csv)
   and load the `prot.csv` file into R.

2. Produce a histogram of
   p-value. [Discuss](http://varianceexplained.org/statistics/interpreting-pvalue-histogram/)
   the distribution of these p-values.

3. Produce a *volcano plot*, showing the log2 fold-change along the x
   axis and -log10 of the adjusted p-value along the y axis. Discuss.

4. Produce an *MA plot*, showing the average intensity (over all
   samples) along the x axis, and the log2 fold-change along the y
   axis. Discuss.

<details>
```{r}
prot <- read_csv("./data/prot.csv")


ggplot(prot, aes(x = p.value)) +
    geom_histogram()

ggplot(prot, aes(x = LogFC,
                 y = -log10(adjp))) +
             geom_point()

ggplot(prot, aes(x = AvgExpr,
                 y = LogFC)) +
             geom_point()

```
</details>
